---
title: "Flexdashboard"
author: "Mehana Leafchild"
date: "11/18/24"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
    theme:
      bootswatch: yeti
      
---
```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(ggthemes)
```

Column {data-width=350}
-----------------------------------------------------------------------


```{r}
np_species <- read_csv("most_visited_nps_species_data_copy.csv") #reading in the data
```


```{r}
specific_parks <- np_species %>%
  filter(ParkName %in% c("Yellowstone National Park",
                         "Grand Canyon National Park",
                         "Joshua Tree National Park"))
```
```{r}
specific_parks <- np_species %>%
  filter(ParkName %in% c("Yellowstone National Park",
                         "Grand Canyon National Park",
                         "Joshua Tree National Park")) #selecing only these specific parks 

filtered_category_data <- specific_parks %>%
  group_by(CategoryName) %>%           # Group by the column of interest
  filter(n() >= 1)      #keep everything over 1

category_counts <- specific_parks %>%
  count(CategoryName, sort = TRUE)      #selecting only categorynames and quantifying

```

In my dashboard I am selecting specific parks to see the variation between the the types of organisms sighted. 

Yellowstone

```{r}
yellowstone_category <- specific_parks %>%
  filter(ParkName == ("Yellowstone National Park")) #selecing only yellowstone park

filtered_yellowstone_category_data <- yellowstone_category %>%
  group_by(CategoryName) %>%           # Group by the column of interest
  filter(n() >= 1)      #keep everything over 1 keep it consistent

yellowstone_category_counts <- yellowstone_category %>%
  count(CategoryName, sort = TRUE)
```

Joshua Tree

```{r}
joshua_tree_category <- specific_parks %>%
  filter(ParkName == ("Joshua Tree National Park")) #selecing only yellowstone park

filtered_joshua_tree_category_data <- joshua_tree_category %>%
  group_by(CategoryName) %>%           # Group by the column of interest
  filter(n() >= 1)      #keep everything over 1 keep it consistent

joshua_tree_category_counts <- joshua_tree_category %>%
  count(CategoryName, sort = TRUE)
```

Grand Canyon

```{r}
grandcanyon_category <- specific_parks %>%
  filter(ParkName == ("Grand Canyon National Park")) #selecing only yellowstone park

filtered_grandcanyon_category_data <- grandcanyon_category %>%
  group_by(CategoryName) %>%           # Group by the column of interest
  filter(n() >= 1)      #keep everything over 1 keep it consistent

grandcanyon_category_counts <- grandcanyon_category %>%
  count(CategoryName, sort = TRUE)
```


The first plot is looking at category observations in all three parks combined.

```{r}
all_parks <- ggplot(category_counts, aes(x = CategoryName, y = n)) +
geom_col() +
coord_flip() + 
  theme_economist_white()
  labs(title = "Number of Species Spotted",
       subtitle = "Within Yellowstone, Joshua Tree, and Grand Canyon",
       x = "Number of Observations",
       y = "Category")

  ggplotly(all_parks)
```

Looking at category observations just in Yellowstone

```{r}
ggplot(yellowstone_category_counts, aes(x = CategoryName, y = n)) +
geom_col() +
coord_flip() + 
  theme_economist() +
  labs(title = "Number of Species Spotted",
       subtitle = "Within Yellowstone",
       x = "Category",
       y = "Percent Number of Observations") +
  scale_y_continuous(labels = scales::percent)
```

Looking at category observations just in Joshua Tree

```{r}
ggplot(joshua_tree_category_counts, aes(x = CategoryName, y = n)) +
geom_col() +
coord_flip() + 
  theme_economist() +
  labs(title = "Number of Species Spotted",
       subtitle = "Within Joshua Tree",
       x = "Number of Observations",
       y = "Category")
```

Looking at category observations just in Grand Canyon

```{r}
grand_canyon <- ggplot(grandcanyon_category_counts, aes(x = CategoryName, y = n)) +
geom_col() +
coord_flip() + 
  theme_economist() +
  labs(title = "Number of Species Spotted",
       subtitle = "Grand Canyon",
       x = "Number of Observations",
       y = "Category")
```

```{r}

grand_canyon + coord_polar()

```